---
  title: "Chapter 3 Homework"
Name: "Curtis Mortensen"
output:
  pdf_document: default
html_document: default
---
  
  # HW Ch 3  {.tabset}
  
  ## Packages used in this key
  ```{r message=FALSE, warning=FALSE}
library(wooldridge)
library(tidyverse)
library(pander)
library(mosaic)
library(foreign)
library(car)
```

# 1 c model_2, 2 a doing two variables, figuring out a certain value, and percentages of the model explained by ____

## Question 1

A problem of interest to health officials (and others) is to determine the effects of smoking during pregnancy on infant health. One measure of infant health is birth weight; a birth weight that is too low can put an infant at risk for contracting various illnesses. Since factors other than cigarette smoking that affect birth weight are likely to be correlated with smoking, we should take those factors into account. For example, higher income generally results in access to better prenatal care, as well as better nutrition for the mother. An equation that recognizes this is
$$ bwght = \beta_{0} + \beta_{1} cigs + \beta_{2} faminc + u $$
  
  #### (a) 
  What is the most likely sign for $\beta{2}$?
  
  
  positive because it's likely to increase the birth weight. 
  
  
  
  #### (b) 
  Do you think $cigs$ and $faminc$ are likely to be correlated? Explain why the correlation might be positive or negative.


Well, I think that's a hard question, and it depends on the years. I think culturally the use of cigarettes has shifted from upper class to mainstream to more lower class over time. So, I suppose nowdays there is likely to be a negative correlation, meaning that increased cigarette use is more likely among lower-earning people. 



#### (c)
How, estimate the equation with and without $faminc$, using the data in "bwght." Report the results in equation form, including the sample size and R-squared. Discuss your results, focusing on whether adding $faminc$ substantially changes the estimated effect of $cigs$ on $bwght$
  
```{r}
data(bwght, package='wooldridge')
faminc <- bwght$faminc
cigs <- bwght$cigs
bwght <- bwght$bwght

m1 <- lm(bwght~cigs)
summary(m1)

m2 <- lm(bwght~cigs+faminc)
summary(m2)

# The first equation is estimated to be y = 119.77 + -0.51 cigs. It has a sample size of ____ and an R-squared of 0.022. The second equation is y = 116.97 +  -0.46 cigs + 0.09 faminc. It has a sample size of ____ and an R-squared of 0.028. However, I do not consider the addition of family income to have a substantial influence on the estimated effect of cigarettes on birth weight. I would base my analysis on the new standard error, not the R-squared. The new standard error is 0.092, and the old standard error was 0.090. So, we haven't removed much of the variance in cigarettes, and I would not consider us to changed the model very much. 
```








## Question 2 
Use the data in "hprice1" to estimate the model
$$ price = \beta_{0} + \beta_{1} sqrft  + \beta_{2} bdrms +u $$
  where $price$ is the household price measured in thousands of dollars.

```{r}
data(hprice1, package='wooldridge')
sqrft <- hprice1$sqrft
bdrms <- hprice1$bdrms
price <- hprice1$price

hprice1 

mp <- lm(price~sqrft+bdrms)
mp


```




#### (a) 
Write out the results in equation form.

y = -19.315 + .1284 sqrft + 1.1982 bdrms




#### (b) 
What is the estimated increase in price for a house with one more bedroom, holding square footage constant?
  
There is an estimated increase of 15.20 in our model, which translates to an increase of $15,200 (rounded). 
  
  
  
  
  
  #### (c) 
  What is the estimated increase in price for a house with an additional bedroom that is 140 square feet in size? Compare this to your answers in part (ii)

```{r}
y = (.1284*140) + 1.1982
y
# When we consider the addition of square feet to our model as well as listing 1 more bedroom the model estimates an increase of $19,200 (rounded). This reveals a flaw in  our previous analysis, because the previous price increase would only be true for a bedroom of 0 sq ft, which of course is impossible. 

summary(mp)

```





#### (d) 
What percentage of the variation in price is explained by square footage and number of bedrooms?
  
If I understand the question correctly, the R-squared value is .62, so our model explains approximately 62% of the variation in our model. Admittedly our intercept helps the model to account for more variation as well. 
  
  
  
  #### (e) 
  The first house in the sample has $sqrft=2438$ and $bdrms = 4$. Find the predicted selling price for this house from the OLS regression line.

```{r}
y1 = -19.315 + (.1284*2438) + (1.1982*4)
y1
# The predicted selling price for this house is $298,500 (rounded) 
```





#### (f) 
The actual selling price of the first house in the sample was $\$300,000$ (so $price=300$). Find the residual for this house. Does it suggest that the buyer underpaid or overpaid for the house?
  
  ```{r}


```






## Question 3
Use the data in "meap93" to answer this question.

#### (a) 
Estimate the model

$$ math10 = \beta_{0} + \beta_{1} \log(expend) + \beta_{2} lnchprg + u $$
  and report the results in the usual form, including the sample size and R-squared. Are the signs of the slope coefficients what you expected? Explain.

```{r}
data(meap93, package='wooldridge')
math10 <- meap93$math10
lnchprg <- meap93$lnchprg
lexpend <- meap93$lexpend
meap93
```






#### (b) 
What do you make of the intercept you estimated in part (i)? In particular, does it make sense to set the two explanatory variables to zero? [*Hint*: Recall that $\log(1)=0$]






#### (c)
Now run the simple regression of $math10$ on $\log(expend)$, and compare the slope coefficients with the estimate obtained in part (i). Is the estimated spending effect now larger or smaller than in part (i)?
  
  
  ```{r}


```






#### (d)
Find the correlation between $$lexpend=\log(expend)$$ and $$lnchprg$$ . Does its sign make sense to you?
  ```{r}

```







#### (e)
Use part (iv) to explain your findings in part (iii)







## Question 4
Use the data "meapsingle" to study the effects of single-parent households on student math performance. These data are for a subset of schools in southeast Michigan for the year 2000. The socio-economic variables are obtained at the ZIP code level (where ZIP code is assigned to schools based on their mailing address)

#### (a)
Run the simple regression of $$math4$$ on $$pctsgle$$ and report the results in the usual format. Interpret the slope coefficient. Does the effect of single parenthood seem large or small?
  
```{r}
data(meapsingle, package='wooldridge')
meapsingle
math4 <- meapsingle$math4
pctsgle <- meapsingle$pctsgle

```







#### (b)
Add the variables $$lmedinc$$ and $$free$$ to the equation. What happens to the coefficient on $$pctsgle$$? Explain what is happening.

```{r}
lmedinc <- meapsingle$lmedinc
free <- meapsingle$free

```





#### (c)
Find the sample correlation between $lmedinc$ and $free$. Does it have the sign you expect?
  ```{r}

```






#### (d)
Does the substantial correlation between $lmedinc$ and $free$ mean that you should drop one from the regression to better estimate the causal effect of single parenthood on student performance? Explain.







#### (e)
Find the variance inflation factors (VIFs) for each of the explanatory variables appearing in the regression in part (ii). Which variable has the largest VIF? Does this knowledge affect the model you would use to study the causal effect of single parenthood on math performance?
  
  ```{r}

```



```{r}

```

