---
title: "Chapter 4 Homework"
Name: "Curtis Mortensen"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(wooldridge)
library(tidyverse)
library(pander)
library(mosaic)
library(foreign)
library(effects)
library(stats)
```

# Homework Chapter 4 {.tabset}

##  Question 1

The following model can be used to study whether campaign expenditures affect election outcomes:

$$ voteA = \beta_0 + \beta_1log(expendA) + \beta_2log(expendB) + \beta_3prtystrA + u $$

where *voteA* is the percentage of the vote received by Candidate A, *expendA* and *expendB* are campaign expenditures by Candidates A and B, and *prtystrA* is a measure of party strength for Candidate A (the percentage of the most recent presidential vote that went to A's party). 


#### a) 

What is the interpretation of $$\beta_0$$?








#### b) 

In terms of the parameters, state the null hypothesis that a 1% increase in A's expenditures is offset by a 1% increase in B's expenditures.







#### c) 

Estimate the given model using the data in VOTE1 and report the results in usual form. Do A's expenditures affect the outcome? What about B's expenditures? Can you use these results to test the hypothesis in part ii?


```{r}

```









#### d) 

Estimate a model that directly gives the t statistic for testing the hypothesis in part ii. What do you conclude? Use a two-sided alternative.

```{r}

```








## Question 2

Use the data in LAWSCH85 for this exercise. 

#### a)

Using the same model as in Problem 4 in Chapter 3, state and test the null hypothesis that the rank of law schools has no ceteris paribus effect on median starting salary. 

In the model,

$$log(salary) = \beta_{0} + \beta_{1}LSAT + \beta_{3}log(libvol) + \beta_{4}log(cost) $$

```{r}

```







#### b) 

Are features of the incoming  class of students- namely, LSAT and GPA-individually or jointly significant for explaining salary? Be sure to account for the missing data on LSAT and GPA. 









#### c) 

Test whether the size of the entering class (*clsize*) or the size of the faculty (*faculty*) needs to be added to this equation; carry out a single test. Be careful to account for missing data on *clsize* and *faculty*. 


```{r}

```






#### d) 

What factors might influence the rank of the law school that are not included in the salary regression?










## Question 3

Use the data in ELEM94_95 to answer this question. The findings can be compared with those in Table 4.1. The dependent variable *lavgsal* is the log of average teacher salary and *bs* is the ratio of average benefits to average salary (by school).

#### a) 

Run the simple regression of *lavgsal* on *bs*. Is the estimated slope statistically different from zero? Is it statistically different from -1? 

```{r}

```








#### b) 

Add the variables *lenrol* and *lstaff* to the regression from part i. What happens to coefficient on *bs* How does the situation compare with that in Table 4.1?

```{r}

```








#### c)

How come the standard error on the *bs* coefficient is smaller in part ii than in part i? Hint: What happens to the error variance versus multicollinearity when *lenrol* and *lstaff* are added?








#### d) 

How come the coefficient on *lstaff* is negative? Is it large in magnitude?










#### e) 

Now add the variable *lunch* to the regression. Holding other factors fixed, are teachers being compensated for teaching students from disadvantaged backgrounds? Explain. 


```{r}

```







#### f) 

Overall, is the pattern of results that you find with ELEM94_95 consistent with the pattern in Table 4.1 (found in the textbook on page 138)?








## Question 4

Use the data in HTV to answer this question. 



#### a) 

Estimate the regression model:

$$ educ = \beta_0 + \beta_1motheduc + \beta_2fatheduc + \beta_3abil + b_4abil^2 + u $$

by OLS and report the results in the usual form. Test the null hypothesis that educ is linearly related to abil against the alternative that the relationship is quadratic. 

```{r}

```






#### b) 

Using the equation in part i, test $$H_0: \beta_1 = \beta_2$$ against a two-sided alternative. What is the p-value of the test?








#### c) 

Add the two college tuition variables to the regression from part i and determine whether they are jointly statistically significant. 









#### d) 

What is the correlation between tuit17 and tuit18? Explain why using the average of the tuition over the two years might be preferred to adding each separately. What happens when you do use the average? 









#### e) 

Do the findings for the average tuition variable in part iv make sense when interpreted causally? What might be going on?














